Reasoning about a Capability Machine with Local Capabilities\\ Provably Safe Stack and Return Pointer Management (without OS Support).

### 2017-02-07 csf-2017/paper.tex
recursive world intuition and a bit technical stuff

\begin{document}

@@

\subsection{Regions}
%Regions intuition
We have three types of regions: $\perma$, $\temp$, and $\revoked$. The
$\perma$ and $\temp$ regions contains a state transition system with
public and private transitions as known from logical relations
from high-level languages\todo{reference}. The state transition
systems can be understood as protocols the memory must follow. The
protocols imposed on memory by $\perma$ regions stay in place
indefinitely. This means that any capability, $\local$ or $\glob$, can
depend on these protocols. The protocols imposed on memory by $\temp$
regions can be revoked at certain points during execution
corresponding to the points when $\local$ capabilities can be
revoked. This means that $\local$ capabilties safely can depend on the
protocols imposed on memory by $\temp$ regions, but $\glob$
capabilities cannot as the $\glob$ capability may stay around
throughout the execution of a program, but the $\temp$ regions may not
stay around.

@@

\begin{theorem}\label{thm:world-existence}
  There exists a \cofe{} $\Wor$ and preorders $\futurestr$ and
  $\futurewk$ such that $(\Wor,\futurestr)$ and $(\Wor,\futurewk)$ are
  preordered \cofes{}, and there exists an isomorphism $\xi$ such that
  {\small
    \begin{align*}              
      \xi : \Wor & \cong \blater (\nats \finparfun \Regions)\\
      \Regions & = \{\revoked\} \uplus \\
                 &\hskip -2cm \{\temp\} \times \States \times \Rels \times (\States \fun (\Wor \monwknefun \UPred{\HeapSegments})) \uplus \\
                 & \hskip -2cm \{\perma\} \times \States \times \Rels \times (\States \fun (\Wor \monstrnefun \UPred{\HeapSegments}))
    \end{align*}
  } and for $W, W' \in \Wor$
  \begin{align*}
    W' \futurestr W & \Leftrightarrow \xi(W') \futurestr \xi(W)   \\
    W' \futurewk W & \Leftrightarrow \xi(W') \futurewk \xi(W)
  \end{align*}
\end{theorem}

\subsection{Worlds}
For this logical relation, we use step-indexed Kripke worlds. They are
defined by the $\Regions$ given in

%Let $\iota.v$ be the projection of the view of a region.
Worlds are finite partial maps from region names to regions,
and region names are simply natural numbers
\begin{align*}
  \Worlds & = \RegionNames \finparfun \Regions\\
 \RegionNames & = \nats.
\end{align*}

>>>>

% domi: IEEEtran.cls sets font size to 10bp, apparently.
% not sure what the difference is with 10pt.
\DeclareMathSizes{10bp}{8}{7}{6}

%\def\IEEEbibitemsep{0pt}

\makeatletter
\setlength\mpr@andskip{1em}
\def\mpr@lineskip{5em}
\def\MathparLineskip{\mpr@lesslineskip}
\makeatother

\newlength{\oldmathindent}
\newenvironment{withmathindent}[1]{\setlength{\oldmathindent}{\mathindent}\setlength{\mathindent}{#1}}{\setlength{\mathindent}{\oldmathindent}}

\begin{document}
\setlength{\mathindent}{.2cm}

@@

\subsection{Worlds}
% Intuition worlds
The worlds are step-indexed Kripke worlds. For the reader not familiar
with these, it is sufficient to think of worlds as finite sets of
regions.

% Regions intuition
We have three types of regions: $\perma$, $\temp$, and $\revoked$. The
$\perma$ and $\temp$ regions contains a state transition system with
public and private transitions as known from logical relations
from high-level languages\todo{reference}. The state transition
systems can be understood as protocols the memory must follow. The
protocols imposed on memory by $\perma$ regions stay in place
indefinitely. This means that any capability, $\local$ or $\glob$, can
depend on these protocols. The protocols imposed on memory by $\temp$
regions can be revoked at certain points during execution
corresponding to the points when $\local$ capabilities can be
revoked. This means that $\local$ capabilties safely can depend on the
protocols imposed on memory by $\temp$ regions, but $\glob$
capabilities cannot as the $\glob$ capability may stay around
throughout the execution of a program, but the $\temp$ regions may not
stay around.

@@

% Recursive worlds intuition
The worlds end up being recursive. The memory can contain capabilities
and capabilities grants authority over part of the memory. A world
models memory, so it also needs to be able to model capabilities in
memory and the protocols on them. Whether a capability satisfies the
protocol of a region may depend on what it grants authority over in
memory which is described by the world.

% Short technical explanation
The recursiveness of the worlds gives rise to a recurisve domain
equation. Theorem~\ref{thm:world-existence} gives the solution to this
recursive domain equation.
\begin{theorem}\label{thm:world-existence}
  There exists a \cofe{} $\Wor$ and preorders $\futurestr$ and
  $\futurewk$ such that $(\Wor,\futurestr)$ and $(\Wor,\futurewk)$ are
  preordered \cofes{}, and there exists an isomorphism $\xi$ such that
  {\small
    \begin{align*}              
      \xi : \Wor & \cong \blater (\nats \finparfun \Regions)\\
      \Regions & = \{\revoked\} \uplus \\
                 &\hskip -2cm \{\temp\} \times \States \times \Rels \times (\States \fun (\Wor \monwknefun \UPred{\HeapSegments})) \uplus \\
                 & \hskip -2cm \{\perma\} \times \States \times \Rels \times (\States \fun (\Wor \monstrnefun \UPred{\HeapSegments}))
    \end{align*}
  } and for $W, W' \in \Wor$
  \begin{align*}
    W' \futurestr W & \Leftrightarrow \xi(W') \futurestr \xi(W)   \\
    W' \futurewk W & \Leftrightarrow \xi(W') \futurewk \xi(W)
  \end{align*}
\end{theorem}
We use the $\Regions$ from Theorem~\ref{thm:world-existence} to define
the worlds:
\begin{align*}
  \Worlds & = \RegionNames \finparfun \Regions\\
 \RegionNames & = \nats
\end{align*}
